\documentclass{article}
\usepackage{color}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{xcolor}
\geometry{letterpaper, total={170mm,257mm}, left=20mm, top=20mm}
\graphicspath{ {./images/} }
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}

\title{MATH2790 Assignments}
\author{Justin Bornais}
\date{\today}

\begin{document}
\maketitle
\section{Assignment 1}
\subsection{Practice Mobius Assignment}

\subsection{Marked Mobius Assignment}
\subsubsection{Question 1}
Find the values of $m$ so that $y=x^m,(x\neq0)$ is a solution of the differential equation  $x^2y''-28xy'+210y=0$

\paragraph{Solution}
$y=xm\Rightarrow y'=mx^{m-1}\Rightarrow y''=m(m-1)x^{m-2}$
\\$x^2y''-28xy'+210y=0$
\\$x^2(m(m-1)x^{m-2})-28x(mx^{m-1})+210(x^m)=0
\\\Rightarrow m(m-1)x^m-28mx^m+210x^m=0
\\\Rightarrow x^m(m(m-1)-28m+210)=0
\\\Rightarrow x^m(m^2-m-28m+210)=0
\\\Rightarrow x^m(m^2-29m+210)=0$
\\\\$m^2-29m+210=0
\Rightarrow (m-14)(m-15)=0
\Rightarrow m=14,15$
\\\\Similarly, $x^m=0\Rightarrow m=0$ but $m\neq0$
\\Thus $m=14$ or $m=15$

\subsubsection{Question 2}
Find the value of $a$ such that the function $y=ax^2-7$ satisfies the differential equation $xy'-y+4(x-1)^2=11-8x$

\paragraph{Solution}
$y=ax^2-7\Rightarrow y'=2ax$
\\$x(2ax)-ax^2+7+4x^2-8x+4-11+8x=0
\\\Rightarrow2ax^2-ax^2+7+4x^2-7=0
\\\Rightarrow ax^2+4x^2=0
\\\Rightarrow (a+4)x^2=0
\\\Rightarrow a+4=0
\\\Rightarrow a=-4$

\subsubsection{Question 3}
Find the value of $c$ such that $y=\frac{1}{x^2+c}$ is a solution to the IVP $y'=-2xy^2;\;y(9)=\frac{1}{15}$

\paragraph{Solution} The condition $y(9)=\frac{1}{15}$ means substituting $x=9$ and $y=\frac{1}{15}$
into the solution $y=\frac{1}{x^2+c}$ gets us $\frac{1}{15}=\frac{1}{9^2+c}\Rightarrow\frac{1}{15}=\frac{1}{81+c}
\\\Rightarrow81+c=15\Rightarrow c=15-81=-66$

\subsubsection{Question 4}
Consider the IVP $(x-1)y'-9y=-8;\;y(1)=1$ Calculate $\frac{\partial f}{\partial y}$

\paragraph{Solution} So $f(x,y)=\frac{-8+9y}{x-1}$ giving $\frac{\partial f}{\partial y}=\frac{9}{x-1}$
\\Both $f$ and $\frac{\partial f}{\partial y}$ are continuous when $x\neq1$.
\\But the existence theorem guarantees that the given IVP has a unique solution on an interval in a region.

\subsubsection{Question 5}
$x^4y'=e^{6x^2+5y^2}$ is or is not separable.

\paragraph{Solution} Well, $\frac{dy}{dx}=\frac{e^{6x^2+5y^2}}{x^4}\Rightarrow \frac{dy}{dx}=\frac{e^{6x^2}}{x^4}\cdot e^{5y^2}$
\\Thus the equation is separable.

\subsubsection{Question 6}
Consider the IVP: $y'-8x^9e^{-y}=0;\;y(0)=8$
\\Find the explicit solution of the IVP.

\paragraph{Solution}
We first use separation of variables to determine the solution to the DE:
\\$y'-8x^9e^{-y}=0\Rightarrow \frac{dy}{dx}=8x^9e^{-y}$
\\$\Rightarrow e^ydy=8x^9dx$
\\$\int e^ydy=\int8x^9dx$
\\$e^y=\frac{8x^{10}}{10}+C$
\\$y=\ln{\left(\frac{8x^{10}}{10}+C\right)}$
\\\\Imposing the initial condition $y(0)=8\Rightarrow \ln{0+C}=8$
\\$C=e^8$
\\\\Thus the solution for the IVP is $y=\ln\left(\frac{8x^{10}}{10}+e^8\right)$

\subsubsection{Question 8}
What is the integrating factor of the linear differential equation? $xy'-15y=x^{16}$ for $x\in(0,\infty)$

\paragraph{Solution}
First, in standard form $\Rightarrow y'-\frac{15}{x}y=x^{15}$
\\Thus $I.F.=e^{\int P(x)dx}=e^{\int -\frac{15}{x}dx}=e^{-15\ln|x|}=e^{(\ln|x|)^{-15}}=\frac{1}{|x|^{15}}=\frac{1}{x^{15}}$ for $x\in(0,\infty)$
   
\subsubsection{Question 9}
Which of the following intervals can be an interval of existence of the solution of the linear differential equation:
$(x-13)y'-12y=(x-12)^3$

\paragraph{Solution}
To standard form: $y'-\frac{12}{x-13}y=\frac{(x-12)^3}{x-13}$ are continuous.
\\Thus the interval of existence is $(13,\infty)$

\subsubsection{Question 10}
Consider the IVP $y'-y=2x;\;y(0)=8$
Find the explicit solution of the IVP.

\paragraph{Solution}
The given equation is a linear equation.
\\$I.F.=e^{\int P(x)dx}=e^{\int -1dx}=e^{-x}$
\\Multiplying by the I.F. we get $e^{-x}y'-e^{-x}y=2xe^{-x}
\Rightarrow\frac{d}{dx}(e^{-x}y)=2xe^{-x}$
\\Integrating both sides W.R.T. x: $e^{-x}y=-2xe^{-x}-2e^{-x}+C\Rightarrow y=-2x-2+Ce^x$
\\Imposing the initial condition $y(0)=8\Rightarrow 8=0-2+C\Rightarrow C=10$
\\Thus the explicit solution to the IVP is $y=-2x-2+10e^x$

%\part{} then \chapter then \section then \subsection then \subsubsection
% then \paragraph then \subparagraph
\end{document}